Time--Frequency Scattering Accurately Models Auditory Similarities Between Instrumental Playing Techniques

### 2020-07-19 paper/paper.tex
page 13

In addition to second-order scattering coefficients (Equation \ref{eq:S2}), we compute joint time--frequency scattering at the first order by convolving the scalogram $\mathbf{U_1 \boldsymbol{x}}$ (Equation \ref{eq:U1}) with the low-pass filter $\boldsymbol{\phi}_T$ over the time dimension, with wavelets $\boldsymbol{\psi}_{\beta}$ ($\beta \geq0$) over the log-frequency dimension, and by applying the complex modulus:
\begin{align}
\mathbf{S_1}\boldsymbol{x}(t, \lambda, \alpha=0, \beta) &=
\big\vert
\mathbf{U_1}\boldsymbol{x} \circledast
\big(\boldsymbol{\phi}_T \otimes \boldsymbol{\psi}_{\beta} \big)
\big\vert (t, \lambda)
\nonumber \\
&=
\left \vert
\iint_{\mathbb{R}^2}
\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
\,\boldsymbol{\phi}_T (t^\prime)
\,\boldsymbol{\psi}_{\beta} (u^\prime)
\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime
\right \vert,
\label{eq:S1}
\end{align}
where the time constant $T$ is the same as in Equation $\ref{eq:S2}$, \ie{} $T=\SI{1000}{\milli\second}$ by default.

Over the time variable, we set the modulation rate $\alpha$ of $\mathbf{S_1}$ to zero in the equation above.
Conversely, over the log-frequency variable, the edge case $\beta=0$ corresponds to replacing the wavelet $\boldsymbol{\psi}_{\beta}$ by the low-pass filter $\boldsymbol{\psi}_F$, comparably to Equation $\ref{eq:Psi}$.
We refer to \cite{anden2019tsp} for more details on the implementation of joint time--frequency scattering.

We adopt the multiindex notation $p = (\lambda, \alpha, \beta)$ as a shorthand for the tuple of frequency, rate, and scale.
The tuple $p$ is known as a scattering path (see \cite{mallat2012cpam}), and may apply to index both first-order ($\mathbf{S_1}$) or second-order ($\mathbf{S_2}$) coefficients.
Given an input waveform $\boldsymbol{x}$, we denote by $\mathbf{S}\boldsymbol{x}$ the feature vector resulting from the concatenation of $\mathbf{S_1}\boldsymbol{x}$ and $\mathbf{S_2}\boldsymbol{x}$:
\begin{equation}
\mathbf{S}\boldsymbol{x}\big(t, p=(\lambda, \alpha, \beta)\big)
=\left\{
\begin{array}{ll}
\mathbf{S_{1}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{if }\alpha=0,\\
\mathbf{S_{2}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{otherwise.}
\end{array}\right.
\end{equation}

An important theorem by Mallat \cite{mallat2012cpam} states that, if $\boldsymbol{y}$ can be obtained from $\boldsymbol{x}$ via the action of some differentiable elastic deformation $\boldsymbol{\tau} \in \mathrm{Diff}(\mathbb{R})$, then the Euclidean norm between $\mathbf{S}\boldsymbol{x}$ and $\mathbf{S}\boldsymbol{y} = \mathbf{S}(\boldsymbol{\tau}\circ\boldsymbol{x})$ can be bounded as follows:
\begin{equation}
\big\Vert \mathbf{S}\boldsymbol{x} - \mathbf{S}(\boldsymbol{\tau}\circ\boldsymbol{x}) \big\Vert^2
\leq
\kappa_{\mathbf{S}} \times
\Vert \boldsymbol{x} \Vert^2 \times
\Vert \boldsymbol{\dot{\tau}} \Vert_{\infty},
\label{eq:lipschitz}
\end{equation}
where the notation $\Vert \boldsymbol{x} \Vert^2$ indicates the squared Euclidean norm, $\Vert \boldsymbol{x} \Vert_{\infty}$ indicates the supremum norm over $\mathbb{R}$, and $\boldsymbol{\dot{\tau}}$ is the derivative of the diffeomorphism $\boldsymbol{\tau}$.
In the equation above, the scalar $\kappa_{\mathbf{S}}$ plays the role of a Lipschitz constant: it depends exclusively on the design choices underlying the scattering operator $\mathbf{S}$ (\eg{} the quality factor $Q$), yet remains independent of the input signal $\boldsymbol{x}$ as well as the diffeomorphism $\boldsymbol{\tau}$ yielding $\boldsymbol{y} = (\boldsymbol{\tau}\circ\boldsymbol{x})$.
In other words, the Euclidean distance between the scattering transform of $\boldsymbol{x}$ and a deformed version of the same signal is bounded from above by the maximal extent of such deformation.

Although the inequality in Equation \ref{eq:lipschitz} was originally formulated for temporal scattering, it recursively generalizes to multivariable scattering operators \cite{oyallon2014iclr}, including joint time--frequency scattering.
In this context, the appeal behind joint time--frequency scattering is that the deformation $\boldsymbol{\tau}$ may be defined in the scalogram domain $\mathbf{U_1}\boldsymbol{x}$, instead of being constrained to the time domain $\boldsymbol{x}$.
Therefore, joint time--frequency scattering coefficients linearize and disentangle small elastic deformations in the time--frequency domain, and represent them activations from a dictionary of multiscale spectrotemporal modulations $\boldsymbol{\Psi}_{\alpha, \beta}$.

>>>>

In addition to second-order scattering coefficients (Equation \ref{eq:S2}), we compute joint time--frequency scattering at the first order by convolving the scalogram $\mathbf{U_1 \boldsymbol{x}}$ (Equation \ref{eq:U1}) with the low-pass filter $\boldsymbol{\phi}_T$ over the time dimension with wavelets $\boldsymbol{\psi}_{\beta}$ ($\beta \geq0$) over the log-frequency dimension, and by applying the complex modulus:
\begin{align}
\mathbf{S_1}\boldsymbol{x}(t, \lambda, \alpha=0, \beta) &=
\big\vert
\mathbf{U_1}\boldsymbol{x} \circledast
\big(\boldsymbol{\phi}_T \otimes \boldsymbol{\psi}_{\beta} \big)
\big\vert (t, \lambda)
\nonumber \\
&=
\left \vert
\iint_{\mathbb{R}^2}
\mathbf{U_1}\boldsymbol{x}(t - t^\prime, \log \lambda - u^\prime)
\,\boldsymbol{\phi}_T (t^\prime)
\,\boldsymbol{\psi}_{\beta} (u^\prime)
\;\mathrm{d}t^\prime \, \mathrm{d}u^\prime
\right \vert,
\label{eq:S1}
\end{align}
where the time constant $T$ is the same as in Equation $\ref{eq:S2}$, \ie{} $T=\SI{1000}{\milli\second}$ by default.

Over the time variable, we set the modulation rate $\alpha$ of $\mathbf{S_1}$ to zero in the equation above.
Conversely, over the log-frequency variable, the edge case $\beta=0$ corresponds to replacing the wavelet $\boldsymbol{\psi}_{\beta}$ by the low-pass filter $\boldsymbol{\phi}_F$.
We refer to \cite{anden2019tsp} for more details on the implementation of joint time--frequency scattering.

We adopt the multi-index notation $p = (\lambda, \alpha, \beta)$ as a shorthand for the tuple of frequency, rate, and scale.
The tuple $p$ is known as a scattering path (see \cite{mallat2012cpam}), and may apply to index both first-order ($\mathbf{S_1}$) and second-order ($\mathbf{S_2}$) coefficients.
Given an input waveform $\boldsymbol{x}$, we denote by $\mathbf{S}\boldsymbol{x}$ the feature vector resulting from the concatenation of $\mathbf{S_1}\boldsymbol{x}$ and $\mathbf{S_2}\boldsymbol{x}$:
\begin{equation}
\mathbf{S}\boldsymbol{x}\big(t, p=(\lambda, \alpha, \beta)\big)
=\left\{
\begin{array}{ll}
\mathbf{S_{1}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{if }\alpha=0,\\
\mathbf{S_{2}}\boldsymbol{x}(t,\lambda,\alpha,\beta) & \textrm{otherwise.}
\end{array}\right.
\end{equation}

%An important theorem by Mallat \cite{mallat2012cpam} states that, if $\boldsymbol{y}$ can be obtained from $\boldsymbol{x}$ via the action of some differentiable elastic deformation $\boldsymbol{\tau} \in \mathrm{Diff}(\mathbb{R})$, then the Euclidean norm between $\mathbf{S}\boldsymbol{x}$ and $\mathbf{S}\boldsymbol{y} = \mathbf{S}(\boldsymbol{x}\circ\boldsymbol{\tau})$ can be bounded as follows:
%\begin{equation}
%\big\Vert \mathbf{S}\boldsymbol{x} - \mathbf{S}(\boldsymbol{\tau}\circ\boldsymbol{x}) \big\Vert^2
%\leq
%\kappa_{\mathbf{S}} \times
%\Vert \boldsymbol{x} \Vert^2 \times
%\Vert \boldsymbol{\dot{\tau}} \Vert_{\infty},
%\label{eq:lipschitz}
%\end{equation}
%where the notation $\Vert \boldsymbol{x} \Vert^2$ indicates the squared Euclidean norm, $\Vert \boldsymbol{x} \Vert_{\infty}$ indicates the supremum norm over $\mathbb{R}$, and $\boldsymbol{\dot{\tau}}$ is the derivative of the diffeomorphism $\boldsymbol{\tau}$.
%In the equation above, the scalar $\kappa_{\mathbf{S}}$ plays the role of a Lipschitz constant: it depends exclusively on the design choices underlying the scattering operator $\mathbf{S}$ (\eg{} the quality factor $Q$), yet remains independent of the input signal $\boldsymbol{x}$ as well as the diffeomorphism $\boldsymbol{\tau}$ yielding $\boldsymbol{y} = (\boldsymbol{\tau}\circ\boldsymbol{x})$.
%In other words, the Euclidean distance between the scattering transform of $\boldsymbol{x}$ and a deformed version of the same signal is bounded from above by the maximal extent of such deformation.
%
%Although the inequality in Equation \ref{eq:lipschitz} was originally formulated for temporal scattering, it recursively generalizes to multivariable scattering operators \cite{oyallon2014iclr}, including joint time--frequency scattering.
%In this context, the appeal behind joint time--frequency scattering is that the deformation $\boldsymbol{\tau}$ may be defined in the scalogram domain $\mathbf{U_1}\boldsymbol{x}$, instead of being constrained to the time domain $\boldsymbol{x}$.
%Therefore, joint time--frequency scattering coefficients linearize and disentangle small elastic deformations in the time--frequency domain, and represent them activations from a dictionary of multiscale spectrotemporal modulations $\boldsymbol{\Psi}_{\alpha, \beta}$.
